---
title: "Data cleaning"
output: html_document
date: "2023-03-13"
---

#getting data
```{r}
library(tidyverse)
library(readxl)
library(stringi)
candy_2015 <- read_excel("raw_data/boing-boing-candy-2015.xlsx")
candy_2016 <- read_excel("raw_data/boing-boing-candy-2016.xlsx")
candy_2017 <- read_excel("raw_data/boing-boing-candy-2017.xlsx")
```
#get column names on data
```{r}
names(candy_2015)
names(candy_2016)
names(candy_2017)
```
# 2015 - deleting columns (based on their names)
```{r}
candy_2015_del_col <- candy_2015 %>%
    select(-c("Please estimate the degrees of separation you have from the following folks [Beyoncé Knowles]",
   "Please estimate the degrees of separation you have from the following folks [Donald Trump]",
 "Please estimate the degrees of separation you have from the following folks [Hillary Clinton]"))
```
## 2015 - deleting colums we do not need (based on their index)
```{r}
candy_2015_del_range <- select(candy_2015_del_col, -c(116:121),
                               -c(100:113),
                               -c(93:95),
                               -c(88, 57, 45, 41, 33, 28, 26, 23, 18, 16, 10))
```
## 2015 - count columns numbers so we can compare
```{r}
ncol(candy_2015_del_col) 
ncol(candy_2015_del_range)
```
## 2016 - deleting colums we do not need (based on their index)
```{r}
candy_2016_del_range <- select(candy_2016, 
                               -c(109:123), 
                               -c(104, 105, 101, 102, 79, 49, 43, 46, 31, 32, 27, 21, 22, 12, 15, 13))
```
## 2016 count columns numbers so we can compare
```{r}

ncol(candy_2016) 
ncol(candy_2016_del_range)
```
## 2017 - deleting colums we do not need (based on their index)
```{r}
candy_2017_del_range <- select(candy_2017, -c(110:120), 
                           -c(107, 108, 104, 105, 81, 43, 31, 26, 27, 15, 13, 114, 21, 12, 49))
```
## 2017 count columns numbers so we can compare
```{r}

ncol(candy_2017) 
ncol(candy_2017_del_range)
```
#obtaining unique names of "country" columns (2016 and 2017, as 2015 doesn't have country data)
```{r}
unique(candy_2017$"Q4: COUNTRY")
unique(candy_2016$"Which country do you live in?")
```
# Changed various names for US, UK and Canada to "United States", "United Kingdom" and "Canada" in 2016
```{r}
candy_2016_del_range_country <- candy_2016_del_range %>%
 mutate(Country = recode( `Which country do you live in?`,
"UK" = "United Kingdom",
"england" = "United Kingdom",
"uk" = "United Kingdom",
"England" = "United Kingdom",
"canada" = "Canada",
"usa" = "United States",
"US" = "United States",
"USA" = "United States",
"United States of America" = "United States",
"uSA" = "United States",
"united states" = "United States",
"United States" = "United States",
"us" = "United States",
"USSA" = "United States",
"U.S.A." = "United States",
"Murica" = "United States",
"USA!" = "United States",
"USA (I think but it's an election year so who can really tell)" = "United States",
"Usa" = "United States",
"U.S." = "United States",
"Us" = "United States",
"Units States" = "United States",
"America" = "United States",
"United states" = "United States",
"USA USA USA" = "United States",
"the best one - usa" = "United States",
"USA! USA! USA!" = "United States",
"u.s." = "United States",
"united states of america" = "United States",
"The Yoo Ess of Aaayyyyyy" = "United States",
"god's country" = "United States",
"USA!!!!!!" = "United States",
"USA! USA!" = "United States",
"United Sates" = "United States",
"Sub-Canadian North America... 'Merica" = "United States",
"Trumpistan" = "United States",
"U.s." = "United States",
"Merica" = "United States",
"UNited States" = "United States",
"United Stetes" = "United States",
"america" = "United States",
"USA USA USA USA" = "United States",
"United  States of America" = "United States",
"United State" = "United States"), .after = `Which country do you live in?`)
``` 
# Changed various names for US, UK and Canada to "United States", "United Kingdom" and "Canada" in 2017
```{r}

candy_2017_del_range_country <- candy_2017_del_range %>%
    mutate(Country = recode(`Q4: COUNTRY`,
"UK" = "United Kingdom",
"england" = "United Kingdom",
"uk" = "United Kingdom",
"England" = "United Kingdom",
"endland" = "United Kingdom",
"U.K" = "United Kingdom",
"canada" = "Canada",
"Can" = "Canada",
"CANADA" = "Canada",
"USA" = "United States",
"us"= "United States",
"usa"= "United States",
"Us"= "United States",
"Us"= "United States",
"US"= "United States",
"Murica"= "United States",
"United States"= "United States",
"united states"= "United States",
"Usa"= "United States",
"United staes"= "United States",
"United States of America"= "United States",
"United states"= "United States",
"u.s.a."= "United States",
"USAUSAUSA"= "United States",
"america"= "United States",
"U.S.A."= "United States",
"unhinged states"= "United States",
"united states of america"= "United States",
"Unites States"= "United States",
"US of A"= "United States",
"The United States"= "United States",
"North Carolina"= "United States",
"Unied States"= "United States",
"U S"= "United States",
"u.s."= "United States",
"The United States of America"= "United States",
"unite states"= "United States",
"U.S."= "United States",
"USA? Hard to tell anymore.."= "United States",
"'merica"= "United States",
"usas"= "United States",
"Pittsburgh"= "United States",
"United State"= "United States",
"New York"= "United States",
"Trumpistan"= "United States",
"United Sates"= "United States",
"USa"= "United States",
"I pretend to be from Canada, but I am really from the United States."= "United States",
"Ahem....Amerca"= "United States",
"United Stated"= "United States",
"New Jersey"= "United States",
"California"= "United States",
"United ststes"= "United States",
"America"= "United States",
"United Statss"= "United States",
"murrika"= "United States",
"USA! USA! USA!"= "United States",
"USAA"= "United States",
"Alaska"= "United States",
"united States"= "United States",
"N. America"= "United States",
"USSA"= "United States",
"u s a"= "United States",
"United Statea"= "United States",
"united ststes"= "United States",
"USA USA USA!!!!"= "United States"),.after = `Q4: COUNTRY`)

```
# Named countries other than UK,USA and Canada as OTHER (2016 and 2017)
```{r}

candy_2016_country_clean <- candy_2016_del_range_country %>%
 mutate(Country = if_else( Country == "United States", "United States",
                    ifelse(Country == "United Kingdom", "United Kingdom",
                    ifelse(Country == "Canada", "Canada", "OTHER"
                    ))))

candy_2017_country_clean <- candy_2017_del_range_country %>%
     mutate(Country = if_else( Country == "United States", "United States",
                    ifelse(Country == "United Kingdom", "United Kingdom",
                    ifelse(Country == "Canada", "Canada", "OTHER"
                    ))))
```
#introducing column "age" in all three datasets (2015, 2016, 2017) and populating it with mutate
```{r}
candy_2015_age <- candy_2015_del_range %>% mutate(age = as.integer(`How old are you?`),.after = `Timestamp`)

candy_2016_age <- candy_2016_country_clean %>% mutate(age = as.integer(`How old are you?`),.after = `Your gender:`)

candy_2017_age <- candy_2017_country_clean %>% mutate(age = as.integer(`Q3: AGE`),.after = `Q2: GENDER`)
```
#running Janitor for 2015 and 2016
```{r}
candy_2015_candy_name <- janitor::clean_names(candy_2015_age)
candy_2016_candy_name <- janitor::clean_names(candy_2016_age)
```
# Renaming candy names column in 2017
```{r}

candy_2017_candy_name_half <- candy_2017_age %>%
rename_with(str_remove,"Q6 \\|", .cols= starts_with("Q6 |"))
candy_2017_candy_name <- janitor::clean_names(candy_2017_candy_name_half)
candy_2017_candy_name
```

## 2015 and 2016: creating a "year" column with mutate and extracting data for it from a timestamp column
```{r}
candy_2015_year <- candy_2015_candy_name %>% 
  mutate(year = format(as.Date(candy_2015_candy_name$timestamp, format="%d/%m/%Y"),"%Y"), .after = "timestamp")

candy_2016_year <- candy_2016_candy_name %>% 
  mutate(year = format(as.Date(candy_2016_candy_name$timestamp, format="%d/%m/%Y"),"%Y"), .after = "timestamp")

```
# 2017 creating a column "year" and populating it with "2017"
```{r}
candy_2017_year <- candy_2017_candy_name %>% 
  mutate(year = "2017", .after = "internal_id")
candy_2017_year
```
#2016 and 2017 - renaming "gender"columns.
```{r}
candy_2016_gender <- candy_2016_year %>% 
  rename("gender" = "your_gender")
candy_2017_gender <- candy_2017_year %>% 
  rename("gender" = "q2_gender")
```
#finding names of columns unique for 2015 and 2016
```{r}
candy_2015_test <- names(candy_2015_year)
candy_2016_test <- names(candy_2016_gender)
intersect(candy_2015_test, candy_2016_test)
setdiff(candy_2015_test, candy_2016_test)
view(candy_2016_test)
```
#unifying candy names in 2015 and 2016
```{r}
candy_2015_year
candy_2016_gender
rename_candy_2016 <- candy_2016_gender %>%
    rename ("bonkers"= "bonkers_the_candy",
            "box_o_raisins" = "boxo_raisins",
            "dark_chocolate_hershey" = "hersheys_dark_chocolate",
            "sweetums" = "sweetums_a_friend_to_diabetes"
            )
```
#finding names of columns unique for 2016 and 2017
```{r}
candy_2017_test <- names(candy_2017_gender)
candy_2016_renamed_test <- names(rename_candy_2016)
setdiff(candy_2017_test, candy_2016_renamed_test)
view(candy_2016_renamed_test)
```
#unifying candy names in 2015 and 2016
```{r}
rename_candy_2017 <- candy_2017_gender %>%
    rename ("bonkers"= "bonkers_the_candy",
            "box_o_raisins" = "boxo_raisins",
            "dark_chocolate_hershey" = "hersheys_dark_chocolate",
            "sweetums" = "sweetums_a_friend_to_diabetes"
                   )
view(rename_candy_2017)
```
#COMBINING ALL 3 DATASETS with bind_rows, creating a dataset named combine_candy
```{r}

combine_candy <- bind_rows(candy_2015_year, rename_candy_2016, rename_candy_2017)
view(combine_candy)
```
#comparing number of columns in 3 datasets and a number of columns in combine_candy dataset
```{r}
ncol(candy_2015_year)
ncol(rename_candy_2016)
ncol(rename_candy_2017)
ncol(combine_candy)
```
#deleting columns we don't need from combine_candy
```{r}
clean_candy <- select(combine_candy, -c (1,4,24,29,33,49,81,82,85,86,87,88,
                                         91,93,95,74,112,113,114,116,117,120,
                                         121, 8, 115))
```
# Unifying column "are you trick or treating" (because we forgot to do it before; as a result, we united three columns and cleaned values in them from NA)
```{r}
clean_candy_final <- clean_candy %>%
    unite("are_you_trick_or_treating", c("are_you_going_actually_going_trick_or_treating_yourself",
           "q1_going_out")) %>%
    mutate(are_you_trick_or_treating = str_remove_all(are_you_trick_or_treating, "_NA")) %>% 
  mutate(are_you_trick_or_treating = str_remove_all(are_you_trick_or_treating, "NA_"))
names(combine_candy)
```
#checking number of columns in a final dataset
```{r}

ncol(clean_candy_final)
```
#writing dataset to a file
```{r}
write_csv(clean_candy_final, "clean_data/clean_candy.csv")
```





